---
title: "ST558-651 Homework 9"
author: "James Carr"
date: "6/29/2021"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(ggplot2)
library(caret)
library(tidyverse)
library(GGally)
```

## Packages
This *.rmd requires knitr, ggplot2, caret, tidyverse, and ggAlly libraries.

## Read in data
```{r read_csv, warning=FALSE, message=FALSE}
#read in data for both and add flag for quality
test <- read_csv('wineQualityTest.csv')
train <- read_csv('wineQualityTrain.csv')
```
In the above code, the data is read in for the test and training sets. A new column is created as a factor for quality of wine greater than six (equal to 1) or less than 6 (equal to 0).

## Summarizing Data
### Pairs
In the below correlation pairs data, 
```{r pair_data, fig.height = 9, fig.width = 15, warning=FALSE, message=FALSE}
ggpairs(train, ggplot2::aes(colour=type))
```

### Summaries
```{r}
red_sum <-  apply(X = train %>% filter(type=='Red') %>% 
                    select(-type), MARGIN = 2,
  			FUN = function(x) { 
  				temp <- c(min(x),
  				          quantile(x, 0.25),
  				          median(x),
  				          mean(x, na.rm = TRUE),
  				          quantile(x, 0.75),
  				          max(x))
  				names(temp) <- c("min", "1st qtr", "median",
  				                  "mean", "3rd Qtr", "max")
  				temp
  			}
  			)

white_sum <-  apply(X = train %>% filter(type=='White') %>% 
                    select(-type), MARGIN = 2,
  			FUN = function(x) { 
  				temp <- c(min(x),
  				          quantile(x, 0.25),
  				          median(x),
  				          mean(x, na.rm = TRUE),
  				          quantile(x, 0.75),
  				          max(x))
  				names(temp) <- c("min", "1st qtr", "median",
  				                  "mean", "3rd Qtr", "max")
  				temp
  			}
  			)

kable(red_sum, digits=1, caption='Red Wine Summary')
kable(white_sum, digits=1, caption='White Wine Summary')
```


## General Linear Models
```{r glm_models}
#full additive model
glm1 <- train(quality ~ ., data=train, method='lm',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))
#full interaction model
glm2 <- train(quality ~ .^2, data=train, method='lm',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))
#simple model (most correlated to quality was alochol)
glm3 <- train(quality ~ alcohol, data=train, method='lm',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))
#reduced model with no interacitons
#just picking a few based on the correlation matrix
glm4 <- train(quality ~ alcohol + type + sulphates + density, 
              data=train, method='lm',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))

#reduced model with interacitons
#just picking a few based on the correlation matrix
glm5 <- train(quality ~ alcohol + sulphates*type + density*type, 
              data=train, method='lm',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))

models <- rbind(glm1$results, glm2$results, glm4$results, 
            glm5$results, glm3$results)
rownames(models) <- c('Full Additive Model',
                      'Full Interaction Model',
                      'Reduced Additive Model',
                      'Reduced Interaction Model',
                      'Simple Model')
kable(models, caption='General Linear Model Results', digits=2)
```
In the above models, the full interaction model showed the best performance with 10-fold cross validation. 

## Logistic Models
```{r log_models}
train.log <- train %>% mutate(qual_fg = if_else(quality < 6, 0, 1)) %>%
                       select(-quality)
train.log$qual_fg <- as.factor(train.log$qual_fg)
test.log <- test %>% mutate(qual_fg = if_else(quality < 6, 0, 1)) %>%
                       select(-quality)
test.log$qual_fg <- as.factor(test.log$qual_fg)


#full additive model
log1 <- train(qual_fg ~ ., data=train.log, method='glm', family='binomial',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))
#full interaction model
log2 <- train(qual_fg ~ .^2, data=train.log, method='glm', family='binomial',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))
#simple model (most correlated to quality was alochol)
log3 <- train(qual_fg ~ alcohol, data=train.log, 
              method='glm', family='binomial',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))
#reduced model with no interactions
#just picking a few based on the correlation matrix
log4 <- train(qual_fg ~ alcohol + type + sulphates + density, 
              data=train.log, method='glm', family='binomial',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))

#reduced model with interactions
#just picking a few based on the correlation matrix
log5 <- train(qual_fg ~ alcohol + sulphates*type + density*type, 
              data=train.log, method='glm', family='binomial',
              preProcess=c('center', 'scale'),
              trControl=trainControl(method='cv', number=10))

models <- rbind(log1$results, log2$results, log4$results, 
            log5$results, log3$results)
rownames(models) <- c('Full Additive Model',
                      'Full Interaction Model',
                      'Reduced Additive Model',
                      'Reduced Interaction Model',
                      'Simple Model')
kable(models, caption='Logistic Regression Model Results', digits=2)
```
In the above models, the additive models seem to match the more complex interaction models so the full additive model will be selected.

## Final Model Performance
```{r test_comparison}
pred1 <- predict(glm2, newdata=test)
rmse <- postResample(pred1, obs=test$quality)[1]
pred2 <- predict(log1, newdata=test.log)
acc <- postResample(pred2, obs=test.log$qual_fg)[1]
scr <- c(rmse, acc)
scr.name <- c('RMSE', 'Accuracy')
mdl.name <- c('Full Interaction Linear Model', 'Full Additive Logistic Model')
tbl <- unname(cbind(mdl.name, scr.name, round(unname(scr), 2)))
colnames(tbl) <- c('Model Name', 'Score Type', 'Score')
kable(tbl, caption='Final Model Scores', digits=2)
```

## Final Regression Model Description
For the regression model, after cross validation I ended up using the full interaction model. It has every term and all of their interactions.
```{r echo=FALSE}
glm2$finalModel
```



## Final Logistic Model Description
For the logistic model, after cross validation I used the full additive model with no interactions. It has every predictor available. 
```{r echo=FALSE}
log1$finalModel
```

