---
title: "ST 558-651 Homework 1"
author: "James Carr"
date: "5/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Subsetting Vectors
```{r}
#creating vector elements 2 to 12 counting by 2
vec <- seq(2, 12, by=2)

#here we print only the odd indexes
vec[c(1, 3, 5)]

#print without first element
vec[-1]

#print without first and third element
vec[c(-1, -3)]

#print all values but the last value
vec[-length(vec)]

#save the vector in reverse
revVec <- rev(vec)

#print first element 5 times
vec[rep(1, 5)]

```

## Subsetting Matrices
```{r}
#matrix w/ 4 rows, 3 cols, of uniform data between 0 and 1
unifMat <- matrix(runif(12, 0, 1),nrow=4, ncol=3)

#odd rows replaced by 2 times value
unifMat[c(1, 3),] <- unifMat[c(1, 3),] * 2

#even rows replaced by 1/2 times value
unifMat[c(2, 4),] <- unifMat[c(2, 4),] * 0.5

#convert vec to matrix and print dimensions
matVec <- as.matrix(vec)
dim(matVec)

#using the drop command to subset
drop_true <- unifMat[1, ,drop=TRUE]
drop_false <- unifMat[1, ,drop=FALSE]
```
Using drop = true seems to return the values as a vector

## Subsetting Data Frames
```{r}
#showing 3 ways to access the 4th column of the data set
names(iris)[4]
iris$Petal.Width
iris[,4]
iris[,"Petal.Width"]

#create extra column whose entrys are all 1s
iris$Extra.Col <- 1

#printing out mtcars columns in sorted order
mtcars <- mtcars[,sort(names(mtcars))]
mtcars

#removing column called Sepal.Width
iris$Sepal.Width <- NULL
names(iris)

#replacing all periods in column name with underscores
names(iris) <- gsub('\\.', '_', names(iris))
names(iris)

```

## Subsetting Lists
```{r}
#creating list 1 and showing 2 ways to access the 2nd element
lst1 <- list(a = vec, b = revVec, c = unifMat)
lst1$b
lst1[[2]]

#creating a 2nd list with list 1 as its only element
lst2 <- list(lst1)
#subsetting list 2 to access b in list 1
lst2[[1]]$b

#test what happens if only using 1 set of square brackets
test <- lst1[1]
test2 <- lst1[[1]]
str(test)
str(test2)
test[1]
test2[1]
```
When using only 1 bracket, it returns the vector as a list of 1 and inside of the list of 1 is the element you want to look at. If you use both brackets, it does not return a list of 1, but it returns the element. 

This could be useful if you want to subset a list to combine with another list, I suppose.
