---
title: "ST558-651 Homework 6"
author: "James Carr"
date: "6/1/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

## Databases
### Local Databases
#### Question 1
*Create connection to google sandbox. Then query a table using dplyr and sql* 
```{r con_goog, warning=FALSE, message=FALSE}
library(bigrquery)
library(DBI)
con <- dbConnect(bigrquery::bigquery(), project = "publicdata",
                 dataset = "samples",
                 billing = "st558-james-carr")
```
A list of tables in the connection to the Google sandbox printed using `dbListTables(con)`: <br>
`r dbListTables(con)`

I think a good exercise would be to determine the number of cigarettes per day by mother birth state (top 10):
```{r dplyr_goog}
natality <- tbl(con, 'natality')
cig_per_day <- natality %>%
                  select(mother_birth_state, cigarettes_per_day) %>%
                  group_by(mother_birth_state) %>%
                  summarise(num_cigs = sum(cigarettes_per_day)) %>%
                  arrange(desc(num_cigs)) 

```
Using the `dplyr` function, here is the number of cigarettes per day summarised:
`r kable(head(cig_per_day, n=10L))`

Now, let's try to get the same results using SQL:
```{r sql_goog}
sql_text <- paste('select top 10 *'
                 ,'from ('
                        ,'select mother_birth_state'
                        ,',sum(cigarettes_per_day) as num_cigs'
                        ,'from natality'
                        ,'group by other_birth_state'
                        ,'order by num_cigs desc)'
                 ,sep=' ')

sql_text <- paste('select mother_birth_state'
                 ,      ',sum(cigarettes_per_day) as num_cigs'
                 ,'from natality'
                 ,'group by mother_birth_state'
                 ,'order by num_cigs desc'
                 ,'limit 10'
                 ,sep=' ')
cig_per_day <- dbGetQuery(con, sql_text)

```
Using `sql`, here is the number of cigarettes per day summarised:
`r kable(cig_per_day)`

The results are exactly the same as the `dplyr` function, which is what we were hoping for.

### Local Database
#### Question 2
*Query the local db to return playlists and playlist_track tables*
*Store them as pl and pl_tracks*
*How many playlists in the database?* 
```{r local_db, warning=FALSE, message=FALSE}
library(RSQLite)
db <- dbConnect(RSQLite::SQLite(), "chinook.db")
db_tbls <- dbListTables(db)
pl <- tbl(db, 'playlists') %>% collect()
pl_tracks <- tbl(db, 'playlist_track') %>% collect()
```
The code above successfully connects to a local database within the working directory. The playlist table contains `r nrow(pl[1])` playlists.

#### Question 3
*Inner join pl and pl_tracks. Freq table on tracks w/ playlist name* 
```{r join}
pl_join <- inner_join(pl, pl_tracks, by='PlaylistId')
pl_tbl <- pl_join %>%
            count(Name) %>%
            rename('Frequency' = 'n') 
```
Using a `dplyr` inner join on PlaylistId yields the following frequency table: 
`r kable(pl_tbl)`

#### Question 4
*Read in tracks and albums and store them* 
```{r tracks}
tracks <- tbl(db, 'tracks') %>% collect()
albums <- tbl(db, 'albums') %>% collect()
```
Tracks and Albums read in and stored. Tracks has `r ncol(tracks)` variables and albums has `r ncol(albums)` variables.

#### Question 5
*Filter pl_join to Grunge (LETS GOOO), left join w/ tracks* 
```{r grunge}
grunge <- pl_join %>% filter(Name == 'Grunge')
grunge_join <- left_join(grunge, tracks, by='TrackId')
```
Below is the table of grunge left joined with tracks. It contains `r ncol(grunge_join)`, since it combines tracks which has `r ncol(tracks)` columns, and grunge which contains `r ncol(grunge)` - and they have 1 column in common which is used to join. It appears all trackIds in grunge have a corresponding track, as there are no nulls. And of course, it has `r nrow(grunge_join)` rows since we took a left join of grunge which has `nrow(grunge)` rows as well. However, both the playlist and the track are called 'Name', leading to a bit of a conflict.
`r (kable(grunge_join))`

#### Question 6
*Rename the conflicting columns* 
```{r rename_cols}
grunge <- grunge %>% rename('PlaylistName' = 'Name')
tracks <- tracks %>% rename('TrackName' = 'Name')
grunge_join <- left_join(grunge, tracks, by='TrackId')
```
After renaming the variables, the conflict no longer exists: 
`r (kable(grunge_join))`

#### Question 7
*Disconnect from the database* 
```{r disco}
dbDisconnect(db) #disconnect from local db
dbDisconnect(con) #disocnnect from server db
```
Disconnected from the local and server database.

## Querying an API
#### Question 1
*Install and load jsonlite and httr* 
```{r load_stuff, warning=FALSE, message=FALSE}
library(httr)
library(jsonlite)
```
Libraries loaded.

#### Question 2
*look at documentation https://pokeapi.co/docs/v2* 
*what are some tables we can access?* 
Tables we can access are:
* Berries
* Berry Firmnesses
* Berry Flavors
* Contest Types
* Contest Effects
* Super Contest Effects
* etc (it appears to be a long list)

#### Question 3
*Pull pikachu attribute from the pokemon table* 
```{r get_pikachu}
base_url <- 'https://pokeapi.co/api/v2'
tbl <- 'pokemon'
pokemon <- 'pikachu'
full_url <- paste0(base_url, '/', tbl, '/', pokemon)
```
After constructing the full url, we should be able to access the API: 
`r full_url`

#### Question 4
*Use get to get data in raw form* 
*Use content to turn into json form* 
*Convert it to a list using fromJSON* 
```{r pikachu_scrape}
poke_raw <- httr::GET(url=full_url)
poke_con <- httr::content(poke_raw)
poke_list <- jsonlite::fromJSON(full_url, flatten=TRUE)
```
Currently, my json list is of length `r length(poke_list)`, so hopefully they've added a new pikachu since this homework was created?!? According to homework documentation, 17 expected.

#### Question 5
*How many moves can pikachu do?* 
The documentation lists `r length(poke_list$moves$move.name)` moves that pikachu can do.

#### Question 6
*Which games/versions is pikachu in?* 
```{r pika_games}
pika_games <- poke_list$game_indices %>%
                select(game_index, version.name) %>%
                group_by(game_index, version.name)
```
Pikachu is apparently in the following games and versions:
`r kable(pika_games, align=c('c', 'c'))`



