---
title: "ST558-651 Homework 8"
author: "James Carr"
date: "6/5/2021"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(parallel)
library(knitr)
```
### Read Data
```{r read_data, warning=FALSE, message=FALSE}
data <- read_csv('concentration.csv') %>%
              filter(variety == 'M.giganteus' |
                     variety == 'S.ravennae')
```
The data was read in successfully using the `read_csv` function. We have `r nrow(data)` rows and `r ncol(data)` columns.

### For Loop
```{r for_loop}
for_loop <- function(data, resp, pred) {
  betas <- cbind(rep(NA, 5000), rep(NA, 5000))
  for (i in 1:nrow(betas)) {
    boot <- sample_n(data, size=nrow(data), replace=TRUE)
    lm.fit <- lm(paste(resp, ' ~ ', pred, " + ", "I(",
                     pred, "^2)",sep = ""),
             data=boot)
    betas[i,] <- summary(lm.fit)$coefficients[,1][2:3]
  }
  max_est <- mean(-betas[,1] / (2*betas[,2]))
  max_se <- sd(-betas[,1] / (2*betas[,2]))
  return(list('mean' = max_est, 'sd' = max_se))
}
for_loop_ests <- for_loop(data, 'Total_lignin', 'concentration')
```
With the for loops, the estimated mean is `r round(for_loop_ests$mean, 1)` and the estimated standard deviation is `r round(for_loop_ests$sd, 1)`.

### Replicate
```{r replicate}
boot_fun <- function(data, resp, pred) {
  boot <- sample_n(data, size=nrow(data), replace=TRUE)
  lm.fit <- lm(paste(resp, ' ~ ', pred, " + ", "I(",
                     pred, "^2)",sep = ""),
             data=boot)
  return (summary(lm.fit)$coefficients[,1][2:3])
}

betas <- replicate(5000, boot_fun(data, 
                                   'Total_lignin', 
                                   'concentration'))


max_est <- mean(-betas[1,] / (2*betas[2,]))
max_se <- sd(-betas[1,] / (2*betas[2,]))
```
Using replicate instead of a for loops, the estimated mean is `r round(max_est, 1)` and the estimated standard deviation is `r round(max_se, 1)`.

### Wrapper Function
```{r wrapper}
seBootFun <- function(data, resp, pred) {
  betas <- replicate(5000, boot_fun(data, 
                                   resp, 
                                   pred))
  max_se <- sd(-betas[1,] / (2*betas[2,]))
  return (sd = max_se)
}


bothBootFun <- function(data, resp, pred) {
  betas <- replicate(5000, boot_fun(data, 
                                   resp, 
                                   pred))
  max_est <- mean(-betas[1,] / (2*betas[2,]))
  max_se <- sd(-betas[1,] / (2*betas[2,]))
  return (list(mean = max_est, sd = max_se))
}
```
Use the wrapper function, we return the estimated standard deviation of `r round(seBootFun(data, 'Glucose', 'concentration'), 1)`


### Parallel Computing
```{r}
response <- c('Total_lignin',
              'Glucose',
              'Xylose',
              'Arabinose')

cluster <- makeCluster(detectCores() - 1)
clusterEvalQ(cluster,library(tidyverse))
clusterExport(cluster,c("seBootFun", "boot_fun"))
par_comp <- parSapply(cl=cluster, FUN=seBootFun, data=data, 
                X=response, pred='concentration')
stopCluster(cluster)

```
Below are the standard deviation results from parallel computing the 4 responses.
`r kable(par_comp, digits=1)`

### Estimates for each

```{r est_all}

cluster <- makeCluster(detectCores() - 1)
clusterEvalQ(cluster,library(tidyverse))
clusterExport(cluster,c("bothBootFun", "boot_fun"))
par_comp <- parSapply(cl=cluster, FUN=bothBootFun, data=data, 
                X=response, pred='concentration')
stopCluster(cluster)
```
Below are the estimates and standard deviations from all. I did use parallel computing again because the functions were all set up and it was easy to do.
`r kable(par_comp, digits=1)`
