---
title: "ST558-651 Homework 4"
author: "James Carr"
date: "5/24/2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

### Number 1
*Create a function called* **count_my_change**
*5 arguments for dollar, qtr, dime, nickel, and penny. all default 0* 
```{r num1}
count_my_change <- function(dollar=0, quarter=0, dime=0, 
                            nickel=0, penny=0) {
  change <- (dollar + quarter * 0.25 + dime * 0.10 + 
          nickel * 0.05 + penny * 0.01)
  str <- paste('Your total is $', change, sep='')
  return (str)
}
count_my_change(1)
count_my_change(1, 1, 1, 1, 1)
count_my_change(1, 2, 3, 4, 5)
```
The function, count_my_change, appears be counting the money and outputting the value to a sentence correctly.

### Number 2
*Explain what the difference is between an infix and prefix function.* <br>
With a prefix conversion, the operator comes before the operands. In an infix, the operator is between the operands.


### Number 3
*Write your own infix function to replace cbind*
```{r num3}
`%add_cols%` <- function(col1, col2) {
  both_cols <- cbind(col1, col2)
  return(both_cols)
}

#the following columns were provided to be tested
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

df1
df2
df1 %add_cols% df2

vec1
df1 %add_cols% df2 %add_cols% vec1
```
It seems to have worked. There are now 5 columns, 2 from each data frame and the 5th from the vector. The vector does not have a name, so it is given the name col2, which is the temporary variable it was passed in. 

### Problem 4
*Create a function that returns the interquartile range*
*Remove name on the returned vector*
*Allow arguments to be passed to quantile function*
*Verify if numeric, confirm either 1 column or a vector*
```{r num4, error=TRUE}
getIQR <- function(vector, ...) {
  #if dataframe with 1 column, extract vector
  #if not a dataframe, and not a vector, then throw error
  if (is.vector(vector)) {
    temp_vec <- vector
  } else if (is.data.frame(vector) & ncol(vector) == 1) {
    if (is_tibble(vector)) {
      temp_vec <- pull(vector[,1]) #extract from tibble
    } else {
      temp_vec <- vector[,1] #extract from data frame
    }
  } else {
    stop('This function only accepts a numeric column of data.')
  }
  
  #if we're here, we have a vector
  #check if numeric
  if (!is.numeric(temp_vec)) {
    stop('This function only accepts a numeric column of data.')
  }
  
  #at this point the variable vector is a numeric vector
  #run quantiles, add options provided by user, remove names, return
  iqr <- unname(quantile(temp_vec, c(0.25, 0.5, 0.75), ...))
  
  #return vector for chaining
  invisible(vector)
  
  #return interquantiles
  return(iqr)
}

tbl1 <- tibble(c(1, 2, 3, 4, 5))
df3 <- data.frame(c(1, 2, 3, 4, 5))
vec2 <- c(1, 2, 3, 'c')
getIQR('a') #error because it's not numeric
getIQR(df1) # should throw error 2 columns
getIQR(df3) #1 column dataframe - should run
getIQR(tbl1) #1 column tibble - should run
getIQR(vec1) #numeric vector - should run
getIQR(vec2) #non-numeric vector, should throw error
```
The tests above appear to have worked. Data frames/tibbles with more than two columns throw an error. Non-numeric data throws an error as well. 

### Number 5
*Read in titanic data and run w/ function created above*
*must work with NAs and tibbles *
```{r num5, error=TRUE}
titanicData <- readr::read_csv("https://www4.stat.ncsu.edu/~post/ST558/datasets/titanic.csv")

#check iqr (don't remove NAs first)
getIQR(titanicData$age) #expecting error due to NAs
getIQR(titanicData$age, na.rm = TRUE) #should run
getIQR(select(titanicData, age), na.rm = TRUE) #should run
getIQR(select(titanicData, fare), na.rm = TRUE) #should run

iqr_tbl <- rbind(getIQR(select(titanicData, age), na.rm = TRUE),
                 getIQR(select(titanicData, fare), na.rm = TRUE))
colnames(iqr_tbl) <- c('25%', '50%', '75%')
rownames(iqr_tbl) <- c('Age', 'Fare')
```
Similarly, the data frames are handled correct, and NAs are appropriately handled when na.rm is passed to the function. The IQR for age and fare are show in this table: <br>
`r kable(iqr_tbl, digits=2)`

### Number 6
*Write a function getCoefVar that returns sample coef of var*
*allow unnamed args to be passed to sd and mean*
*check named args passed are valid*
```{r num6, error=TRUE}
getCoefVar <- function(data, ...) {
  #if dataframe with 1 column, extract vector
  #if not a dataframe, and not a vector, then throw error
  if (is.vector(data)) {
    temp_vec <- data
  } else if (is.data.frame(data) & ncol(data) == 1) {
    if (is_tibble(data)) {
      temp_vec <- pull(data[,1]) #extract from tibble
    } else {
      temp_vec <- data[,1] #extract from data frame
    }
  } else {
    stop('This function only accepts a numeric column of data.')
  }
  
  #if we're here, we have a vector
  #check if numeric
  if (!is.numeric(temp_vec)) {
    stop('This function only accepts a numeric column of data.')
  }
  
  #at this point the variable vector is a numeric vector
  #get coef of variation
  #add options provided by user
  coef <- sd(temp_vec, ...) / abs(mean(temp_vec, ...)) 
  
  #return data for chaining
  invisible(data)
  
  #return interquantiles
  return(coef)
}

#testing
getCoefVar('a') #error because it's not numeric
getCoefVar(df1) # should throw error 2 columns
getCoefVar(df3) #1 column dataframe - should run
getCoefVar(tbl1) #1 column tibble - should run
getCoefVar(vec1) #numeric vector - should run
getCoefVar(vec2) #non-numeric vector, should throw error

```
This function was tested with similar methods as the previous function. All tests passed.

### Number 7
*Use on same titanic data again*
*verify NA and tibble handling*
```{r num7, error=TRUE}
#check coefVar (don't remove NAs first)
getCoefVar(titanicData$age) #expecting NA due to NAs
getCoefVar(titanicData$age, na.rm = TRUE) #should run
getCoefVar(select(titanicData, age), na.rm = TRUE) #should run
getCoefVar(select(titanicData, fare), na.rm = TRUE) #should run
```
NAs appropriately handled. Age has a Coefficient of Variation of `r round(getCoefVar(select(titanicData, age), na.rm = TRUE), 2)` and fare's Coefficient of Variation is `r round(getCoefVar(select(titanicData, fare), na.rm = TRUE), 2)`.

### Number 8
*Create a function that returns both coefficient of variation and the IQR*
```{r num8}
getSummaries <- function(data, ...) {
  #get iqr
  iqr <- getIQR(data, ...)
  #get Coefficient of Variation
  coefVar <- getCoefVar(data, ...)
  
  #return data for chaining
  invisible(data)
  
  #return in list form for proper naming
  return (list(iqr = iqr, coefVar = coefVar))
}
```
The function uses the previously created functions and returns the IQR and Coefficient of Variation in a list (so the list has two objects).

### Number 9
*Run the 2 columns of titanic data through the new function*
```{r num9}
age <- getSummaries(titanicData$age, na.rm = TRUE)
fare <- getSummaries(titanicData$fare, na.rm = TRUE)

summaries <- rbind(c(age$iqr, age$coefVar),
                   c(fare$iqr, fare$coefVar))
colnames(summaries) <- c('25%', '50%', '75%', 'Coef')
rownames(summaries) <- c('Age', 'Fare')

```
Now, with the combined function, we get the following values for age and fare:
`r kable(summaries, digits=2)`
These match the previous values.

### Number 10
*Why can you name a value sd in a function and not cause any issues with the sd function* <br>
Our functions are stored in the global environment, and the variables within them are temporary. The variable sd is stored in the base environment. 